---
title: "Human PBMC profiled on different technologies"
format: html
---

## Dataset

Data from "[Systematic comparison of single-cell and single-nucleus RNA-sequencing methods](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7289686/)", which is a dataset of human PBMC profiled with seven different technologies. The data is available as part of the [SeuratData](https://github.com/satijalab/seurat-data) package. 

```{r init}
library(Seurat)
library(SeuratData)
```

Install the dataset.

```{r installdata, include=FALSE, eval=TRUE}
options(timeout = 1000)
InstallData("pbmcsca")
```

Load in the pbmc systematic comparative analysis dataset.

```{r load, message=FALSE}
seurat_obj <- LoadData("pbmcsca")
seurat_obj
```

## Metadata

Number of cells per technology.

```{r table_of_methods}
table(seurat_obj@meta.data$Method)
```

## Assays

Seurat v5 uses the Assay5 class. Assay5 objects are more flexible, and can be used to store only a data layer, with no counts data. This can be used to create Seurat objects that require less space

```{r getassays}
sapply(Assays(seurat_obj), \(x) class(seurat_obj[[x]]))
```

## Layers in the Seurat v5 object

Seurat v5 assays store data in layers. These layers can store raw, un-normalized counts (`layer='counts'`), normalized data (`layer='data'`), or z-scored/variance-stabilized data (`layer='scale.data'`).

Extract [layer](https://satijalab.org/seurat/articles/seurat5_essential_commands#working-with-layers) names:

```{r layer_names}
Layers(seurat_obj[["RNA"]])
```

We can split the layers into groups. The object contains data from nine different batches (stored in the `Method` column in the object metadata), representing seven different technologies.

```{r splitassay}
seurat_obj[["RNA"]] <- split(seurat_obj[["RNA"]], f = seurat_obj$Method)
Layers(seurat_obj[["RNA"]])
```
After splitting, there are now 18 layers (a `counts` and `data` layer for each batch).

Use `JoinLayers` to join them back up.

```{r joinlayers}
seurat_obj[["RNA"]] <- JoinLayers(seurat_obj[["RNA"]])
Layers(seurat_obj[["RNA"]])
```

## Session info

```{r}
sessionInfo()
```
